### 2024-02-27 paper.tex
Initial commit

\documentclass[12pt, a4paper]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subfigure}

\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\begin{document}
\title{When two-point correlation function fails to describe a structure...}
\author{Foo Bar}
\maketitle

\section{Introduction}
AAA BBB

\section{All the math}
\subsection{Definitions}
Suppose $A$ is some region in a Euclidean space $\mathbb{R}^n$. Let
$A_1, A_2, \dots, A_N$ be subsets of $A$ so that
\begin{align*}
  \cup_{i=1}^N A_i &= A \\
  A_i \cap A_j &= \emptyset, \ i \ne j
\end{align*}
It is said that the point $x$ is in the phase $i$ if $x \in A_i$.

Define a set of characteristic functions $\chi_i(x)$ as follows:
\begin{equation*}
  \chi_i(x) = \left\{
  \begin{array}{ll}
    1 & \quad \text{if $x \in A_i$} \\
    0 & \quad \text{otherwise}
  \end{array}
  \right.
\end{equation*}

The two-point correlation function
$S_2^{(i)}: \mathbb{R}^n \rightarrow \mathbb{R}$ for homogeneous media can
defined in the terms of $\chi$:
\begin{equation}
  S_2^{(i)}(r) = \langle \chi_i(x) \chi_i(x+r) \rangle
  \label{eq:s2}
\end{equation}
where $\langle \dots \rangle$ means ensemble average. Its meaning is given in
the following definition:
\begin{definition}
  The two-point correlation function $S_2^{(i)}(r)$ is a probability that a
  line segment $r$ thrown randomly into $A$ has its both ends in the phase $i$.
  \label{def:s2}
\end{definition}

Two-phase media in which the region $A$ is subdivided in two adjacent phases
(usually called ``void'' and ``solid'') are of special interest for practical
applications. In this case function $S_2^{(1)}$ and $S_2^{(0)}$ has a linear
dependency \cite{Torquato_book}:
\begin{equation*}
  S_2^{(1)}(r) = S_2^{(0)}(r) + 1 - 2S_2^{(0)}(0)
\end{equation*}
In our paper we will stick to $S_2^{(1)}$ function for simplicity.

In practice the region $A$ is a $n$-dimensional array, vectors $r$ and $x$ can
have integer coefficients and $x \in A^{(1)}$ is equivalent to $A[x] = 1$
(element at the index $x$ is equal to $1$). The two-point correlation function
is then computed using fast Fourier transform.

A problem of reconstruction of $A$, i.e. properly finding its subset $A_1$, from
its two-point correlation function is often arises in porous materials
science. In this paper we approach the problem of reconstruction having a
so-called full correlation map of $A$, i.e. the values $S_2^{(1)}(x)$ for
$x \in (1 \dots 2D_1-1)\times(1 \dots 2D_2-1)\times\dots$ where
$D_1, D_2, \dots$ are dimensions of the array $A$. Here multiplication by two
and subtraction of one means that we consider two-point function calculated with
zero padding, described in the next subsection.

\subsection{Boundary conditions}
The two-point correlation function can be calculated with periodic boundary
conditions and zero padding depending on whether a line segment in
\cref{def:s2} is allowed to cross the boundary of array or not. If it is allowed
to do so, its coordinates are ``wrapped'' around the boundary, i.e. taken
$\mod D$, $D = (D_1, D_2, \dots)$ (\cref{fig:s2-computation}).
\begin{figure*}[tp]
  \centering
  \subfigure[Periodic boundary conditions]{
    \includegraphics[width=0.4\linewidth]{images/periodic.png}
    \label{fig:s2-periodic}}
  \hfill
  \subfigure[Zero padding]{
    \includegraphics[width=0.4\linewidth]{images/zeros.png}
    \label{fig:s2-zeros}}
  \caption[]{Computation of the two-point correlation function at the point $2$
    for one-dimensional sequence of length 8.}
  \label{fig:s2-computation}
\end{figure*}

In this paper we will primarily discuss the two-point function calculated with
zero padding.

\subsection{Collisions in the two-point function for sequences of 0's and 1's}
When reconstructing a two-phase porous medium from its correlation function, it
is important to know to what extent a reconstruction will resemble the original
image. Consider the following definition:
\begin{definition}
  Arrays $A$ and $B$ of the same dimensionality have a collision if they have
  the same values of the two-point correlation function in all points of its
  definition, but $A \ne B$.
\end{definition}

Turns out that every array $A$ has a collision of two specific forms, which we
will discuss after describing of a mathematical framework which will help us in
the analysis.

\begin{definition}
  A Z-transform of a sequence $\{a_0, a_1, \dots, a_n\}$ is a polynomial
  $\sum_{i=0}^n a_i z^{-i}$.
\end{definition}
Z-transform is much like the discrete Fourier transform, particulary it has the
same convolution theorem:
\begin{theorem}
  If $g(z)$ and $h(z)$ are images of sequences $a$ and $b$ under Z-transform,
  then $g(z)h(z) = f(z)$ where $f(z)$ is an image of $a*b$ under Z-transform,
  where $a*b$ is a discrete convolution of $a$ and $b$:
  \begin{equation*}
    (a*b)_k = \sum_{i=0}^{n} a_i b_{k-i}
  \end{equation*}
  \label{th:conv}
\end{theorem}
In the case of one-dimensional two-phase porous media $a$ the definition
\ref{eq:s2} can be rewritten as follows
\begin{equation*}
  S_2(r) = \frac{\sum_{i}a_ia_{i+r}}{N(r)}
\end{equation*}
where $N(r)$ is a normalization factor which does not depend on $a$ and can be
omitted for simplicity. The numerator is a discrete autocorrelation which is
equal to a convolution of $a$ with ``reversed'' $a$, i.e. $i$-th element of $a$
becomes $-i$-th element of a reversed sequence. Let $f(z)$ be an image of $a$
under Z-transform. Then an image $s_2(z)$ of the sum above can be calculated
using \cref{th:conv}.
\begin{equation}
  s_2(z) = f(z)f(z^{-1})
  \label{eq:s2z}
\end{equation}
Making a substitution $t^{-1} = z + z^{-1}$ in inverting $s_2(t)$ we will get a
sequence $S_2(r) = (S_2(0), S_2(1), S_2(2), \dots)$ up to a normalization by
$N(r)$. The meaning of this sequence is a number of positive elements in $a$
distanced by $r$.
\begin{example}
  Let $a$ be a sequence $(1, 0, 1, 0, 0, 1)$. Its Z-transform is
  $f(z) = 1 + z^{-2} + z^{-5}$.
  \begin{align*}
    s_2(z) &= f(z)f(z^{-1}) \\
    &= (1 + z^{-2} + z^{-5})(1 + z^2 + z^5) \\
    &= (1 + z^2 + z^5) + (z^{-2} + 1 + z^3) + (z^{-5} + z^{-3} + 1) \\
    &= 3 + (z^2 + z^{-2}) + (z^3 + z^{-3}) + (z^5 + z^{-5})
  \end{align*}
  Thus, we have $s(t) = 3 + t^{-2} + t^{-3} + t^{-5}$. Inverting, we obtain a
  sequence $(3, 0, 1, 1, 0, 1)$, which means that $a$ has three elements in
  phase $1$, zero adjacent elements in phase $1$, one pair which is distanced by
  2 (elements with indices 0 and 2), one pair distanced by 3 (indices 2 and 5)
  and one pair distanced by 5 (indices 0 and 5).
\end{example}

By looking at \cref{eq:s2z} it becomes clear that the following two
substitutions, designated $S_k$ and $R$, do not change $s_2(z)$ and hence, the
two-point correlation function:
\begin{equation}
  \begin{aligned}
    R&: f(z) \rightarrow f(z^{-1}) \\
    S_k&: f(z) \rightarrow z^{-k}f(z)
  \end{aligned}
  \label{eq:group-elts}
\end{equation}
The set
\begin{equation}
  G = \{R, S_0, S_1, S_{-1}, \dots \}
  \label{eq:group}
\end{equation}
and a binary operation $\circ$ (composition) form a group of transforms of $a$
which do not change the two-point function. Its neutral element is $S_0$ and the
associativity and invertibility are trivial to proof.

The transform $R$ means a reversion of a sequence, i.e. writing it ``the other
way around''. The transform $S_k$ means shifting a sequence to the right by $k$
elements. These are trivial transforms which are already present in the
literature. It is clear that any two sequences $a$ and $Ta$, $T \in G$ form a
collision, so if we search for collisions, it is important to ``discard''
trivial collisions like that.

To find non-trivial collisions we firstly introduce a change of variables
$x = z^{-1}$ which is justified by a symmetry of \cref{eq:s2z}. Secondly we
introduce a notion of a reverse of a polynomial and palindrome polynomials.
\begin{definition}
  A reverse of a polynomial $f(x)$ is a polynomial $x^{\deg f}f(x^{-1})$. A
  polynomial is called a palindrome if it is equal to its reverse.
\end{definition}
Examples of palindrome polynomials include $x$, $x + 1$, $x^2 - 2x + 1$, but not
$x^3+x+1$.

Now a mechanism for finding collisions for a two-phase sequence $a$ is
described. Suppose $f(x)$ is a Z-transform of $a$ with aforementioned change of
variables. Its coefficients is therefore belong to a set $\{0, 1\}$ with the
leading and the free coefficients being equal to $1$. The assumption on the free
coefficient is taken without a loss of generality. Indeed, if the free
coefficient is zero, then a monomial $x^n$ divides $f$, and the following
construction can be applied to $\hat{f} = f / x^n$ which has the same
correlation function as $f$.

A reverse of $f(x)$ has the same degree as $f$. Suppose $f$ factors into $M$
different factors in $\mathbb{Z}[x]$:
\begin{equation*}
  f(x) = \prod_{i=1}^M f_i(x)
\end{equation*}
The image of the two-point function is then
\begin{equation*}
  s(x) = f(x)f(x^{-1}) = \prod_{i=1}^M f_i(x) \prod_{i=1}^M f_i(x^{-1})
\end{equation*}
Let now select $1 \le L < M$ non-palindrome factors
$f_{r_1}, f_{r_2}, \dots, f_{r_L}$ of $f$ and replace them with their
reverses. Obviously, the resulting correlation function won't change:
\begin{equation*}
  s(x) = \underbrace{\prod_{i=1}^L x^{\deg f_{r_i}} f_{r_i}(x^{-1}) \prod_{i=1}^{M-L} f_{o_i}(x)}_{\tilde{f}(x)}
  \underbrace{\prod_{i=1}^L x^{-\deg f_{r_i}} f_{r_i}(x) \prod_{i=1}^{M-L} f_{o_i}(x^{-1})}_{\tilde{f}(x^{-1})}
\end{equation*}
where $o_1, o_2, \dots, o_{M-L}$ run through indices of original (not replaced)
factors. Now recombining factors into $\tilde{f}(x)$ and $\tilde{f}(x^{-1})$ we
obtain a polynomial $\tilde{f}$ with the same degree as $f$ and $f(0) = \tilde{f}(0)$.
Now note that with this additional demand on the free coefficient, we can
restrict the group of transforms \cref{eq:group} to a subgroup $\{S_0, R\}$
because a non-zero shift will change the degree of $f$ which is not allowed. We
have two possible outcomes regarding $\tilde{f}$:
\begin{itemize}
\item Only one factor in $f_{r_1}, \dots, f_{r_L}$ is not a palindrome. This
  gives us $\tilde{f}(x) = f(x^{-1})$ which is a trivial collision.
\item At least two factors in $f_{r_1}, \dots, f_{r_L}$ are not palindromes. In
  this case $\tilde{f} \ne f$ and cannot be obtained from $f$ using transforms
  in the subgroup $\{S_0, R\}$.
\end{itemize}

\textcolor{blue}{It is trivial to prove that $\sum_if_i^2 = \sum_i\tilde{f}_i^2$
  where $f_i$ and $\tilde{f}_i$ are coefficients of $f$ and $\tilde{f}$
  respectively. Indeed, $f$ and $\tilde{f}$ have the same correlation function,
  and evaluating $S_2(0)$ gives us the desired result. We also make a conjecture
  that $f$ and $\tilde{f}$ have the same number of non-zero coefficients and
  thus, all coefficients of $\tilde{f}$ are either $0$ or $1$.}
\textcolor{red}{It's better to find a proof for this claim}.

Now it must be clear how to find all the non-trivial collisions of $f$. To do so
one must follow \cref{alg:collisions}.
\begin{algorithm}
  \caption{An algorithm for finding non-trivial collisions of $f$ in the
    two-point correlation function.}
  \label{alg:collisions}
  \begin{algorithmic}[1]
    \Procedure{$Collisions$}{$f$}
    \State $fs \gets factor(f) \in \mathbb{Z}[x]$
    \State $ps \gets filter(isPalindrome, fs)$
    \State $nps \gets fs \setminus d$
    \If{$length(nps) \le 1$}
    \State \textbf{return} $f$ has only trivial collisions
    \EndIf
    \State $(np:nps') \gets nps$
    \State $f' \gets np \times \prod ps$
    \State $cs \gets \text{all possible combinations of $nps'$ or reverses}$
    \State \textbf{return} $map(f'\times, cs)$
    \EndProcedure
  \end{algorithmic}
\end{algorithm}
Here ``all possible combinations of $nps'$ or reverses'' needs
clarification. For each polynomial in $nps'$ either that polynomial or its
reverse is taken in a combination of factors $c$ and $cs$ is a list of all
possible $c$'s. If $f$ has $L$ non-palindrome factors as before, we get
$2^{L-1}$ polynomials (including $f$) which have the same two-point function as
$f$.

It is clear to see, that non-trivial collisions of $f$ found using
\cref{alg:collisions} are indeed all non-trivial collisions of $f$, because the
ring $\mathbb{Z}[x, x^{-1}]$ of images of the two-point function under
Z-transform is a UFD.

\begin{example}
  Let $f = x^{20} + x^{15} + x^{13} + x^{11} + x^{10} + x^7 + x^4 + x^3 + 1 \in \mathbb{Z}[x]$.
  Factors of $f$ are $f_1 = x^{13} + x^{12} + x^{11} + x^{10} + x^4 + x^3 + x^2 + x + 1$
  and $f_2 = x^7 - x^6 + x^3 - x + 1$. We see that these factors are not
  palindromes:
  $S_{-13}R f_1 = x^{13} + x^{12} + x^{11} + x^{10} + x^9 + x^3 + x^2 + x + 1 \ne f_1$ 
  and $S_{-7}R f_2 = x^7 - x^6 + x^4 - x + 1 \ne f_2$. Hence we have 2 polynomials
  which form a non-trivial collision: $f$ and
  $g = x^7 f_1(x) f_2(x^{-1}) = x^{20} + x^{17} + x^{15} + x^{11} + x^{10} + x^8 + x^7 + x^4 + 1$.
  The two-point correlation function of these polynomials is
  $s_2(t) = 9 + 2t + 2t^2 + 4t^3 + 4t^4 + 2t^5 + 2t^6 + 4t^7 + 2t^8 + 2t^9 +
  3t^{10} + 2t^{11} + t^{12} + 2t^{13} + x^{15} + x^{16} + x^{17} + x^{20}$.
  Sequences corresponding to $f$ and $g$ are
  \begin{align*}
    & 100110010011010100001 \quad \text{and} \\
    & 100010011011000101001
  \end{align*}
  Two-point correlation function (without normalization) is
  \begin{equation*}
    (9, 2, 2, 4, 4, 2, 2, 4, 2, 2, 3, 2, 1, 2, 0, 1, 1, 1, 0, 0, 1)
  \end{equation*}
\end{example}

\subsection{Multidimensional case}
All said above works also for the multidimensional case: suppose we have
$n$-dimensional array of bits $A$, i.e. values ``0'' and ``1''. Using
Z-transform we associate a polynomial $f$ with $A$:
\begin{equation*}
  f(z_1, \dots, z_n) = \sum_{i_1=0}^{D_1 - 1} \dots \sum_{i_n=0}^{D_2 - 1}
  a_{i_1\dots i_n} z_1^{-i_1} \dots z_n^{-i_n}
\end{equation*}
where $D_1 \times \dots \times D_n$ is the dimensionality of $A$. Then an image
of the two-point function under Z-transform is
\begin{equation}
  s_2(z_1, \dots, z_n) = f(z_1, \dots, z_n)f(z_1^{-1}, \dots, z_n^{-n})
  \label{eq:s2z-md}
\end{equation}

The same group \cref{eq:group} of transforms \cref{eq:group-elts} gives rise to
trivial collisions in the two-point function, where
\begin{equation}
  \begin{aligned}
    R&: f(z_1, \dots, z_n) \rightarrow f(z_1^{-1}, \dots, z_n^{-1}) \\
    S_{k_1, \dots, k_n}&: f(z) \rightarrow z_1^{-k_1}\dots z_n^{-k_n}f(z)
  \end{aligned}
  \label{eq:group-elts-md}
\end{equation}
where $R$ means a ``reversion'' of $A$ and $S_{k_1, \dots, k_n}$ translates $A$
by a vector $(k_1, \dots, k_n)$. Now, adopting the definition of a palindrome to
the multidimensional case we can apply a change in variables $x_i = z_i^{-1}$
and the \cref{alg:collisions} to finding non-trivial collisions.

\begin{example}
  Consider a two-phase two-dimensional image \cref{fig:2d-example}. Its image
  under the Z-transform is
  $f(z_1, z_2) = z_1^{-1} + z_1^{-2} + z_2^{-1} + z_1^{-1}z_2^{-1} +
  z_1^{-2}z_2^{-1} + z_1^{-1}z_2^{-2}$. An image of the two-point correlation
  function is $s_2(t_1, t_2) = 6 + 3t_1 + t_1^2 + 3t_2 + t_2^2 + \dots$. Only
  the terms contributing to the ``axial'' two-point function are shown here for
  the sake of brevity.
\end{example}
\begin{figure*}[tp]
  \centering
  \includegraphics[width=0.4\linewidth]{images/example-s2-image.png}
  \hfill
  \includegraphics[width=0.4\linewidth]{images/example-s2.png}
  \caption[]{The unnormalized two-point function (right) calculated for a
    two-phase two-dimensional image (left). ``Ones'' are yellow, ``zeros'' are
    purple.}
  \label{fig:2d-example}
\end{figure*}

\subsection{Periodic boundary conditions}
Formula \ref{eq:s2z} can be applied to calculate the two-point function for a
sequence of the length $D$ under periodic boundary conditions if the calculation
if performed modulo $z^D - 1$. The group of transforms \cref{eq:group} does not
change the two-point function. This time, $S_k$ must be understood as a circular
shift by $k$ elements to the right.
\begin{example}
  Consider a sequence $(0, 1, 1, 1, 0, 1, 1)$ of length $7$. The image of its
  correlation function under Z-transform is
  $s_2(z) = 5 + 3(z+z^{-1}) + 2(z^2 + z^{-2}) + 2(z^3 + z^{-3}) + 2(z^4 +
  z^{-4}) + (z^5 + z^{-5})$.
  Taking $s_2(z) \mod x^7-1$ we obtain
  $5 + 3z + 3z^6 + 2z^2 + 2z^5 + 2z^3 + 2z^4 + 2z^4 + 2z^3 + z^5 + z^2 \mod z^7
  - 1 = 5 + 3z + 3z^2 + 4z^3 + 4z^4 + 3z^5 + 3z^6$. The two-point correlation
  function is then $(5, 3, 3, 4, 4, 3, 3)$.
\end{example}

Unfortunately, there is no known way to test a sequence for non-trivial
collisions in periodic boundary conditions.

\section{Probability of encountering a polynomial with non-trivial collision.}

Suppose $f \in \mathbb{Z}$ is a random polynomial with independent uniformly
distributed random coefficients in $\{0, 1\}$ with the free coefficient being
equal to $1$. It is known, that the probability of $f$ being irreducible tends
to $1$ with $d = \deg f \to \infty$ \cite{konyagin1999number}.

Breuillard et~al. \cite{breuillard2019irreducibility} give the following
approximation for this probability based on extended Riemann hypothesis:
\begin{equation}
  P(\text{$f$ is irreducible}) = 1 - \sqrt{\frac{2}{\pi d}} + O(d^{-1})
  \label{eq:prob-irr}
\end{equation}

\textcolor{red}{How to cite a theorem in the paper?}
Another useful result is that for large enough $d$ the probability
of $f$ having factorization $\Phi \tilde{f}$ where $\tilde{f}$ is an irreducible
polynomial and $\Phi$ is a product of cyclotomic polynomials is at least
\begin{equation}
  p = 1 - 2\exp(-c(\log d)^{\beta - 2})
  \label{eq:prob-cyc-1}
\end{equation}
for some $c > 0$ and $\beta > 3$ if $\zeta_K$ has no zeros $\rho$ with
$|1 - \rho| < (\log d)^\alpha/d$, $\alpha > \beta$ for all $K = \mathbb{Q}(a)$
for any root $a$ of $f$.

A stronger bound of the probability of the same factorization with higher
reliance on the Riemann hypothesis is given by:
\begin{equation}
  p = 1 - \exp(-c \sqrt{d} / \log d)
  \label{eq:prob-cyc-2}
\end{equation}
if the Riemann hypothesis holds for $\zeta_k$ for $K = \mathbb{Q}$ and all
fields $K = \mathbb{Q}(a)$ where $a$ is a root of $f$.

Now we will prove a simple lemma which connects these results with collisions in
correlation functions.
\begin{lemma}
  Let $A$ be a sufficiently long random sequence of 0's and 1's where each
  element is chosen randomly and independently with equal probability $1/2$ with
  exception of the first and the last elements which are 1's. Then, with
  probability at least
  \begin{equation*}
    p = 1 - 2\exp(-c(\log d)^\beta)
  \end{equation*}
  for some $c > 0$, $\beta > 1$ the sequence $A$ of the length $d+1$ does not
  have non-trivial collisions in the two-point correlation function.
\end{lemma}
\begin{proof}
  Let $f(x)$ be an image of $A$ under Z-transform and a change of variables
  $x = z^{-1}$. With probability at least $p$ it factors in $\mathbb{Z}[x]$ as
  $\Phi \tilde{f}$, where $\Phi$ is a product of cyclotomic polynomials and
  $\tilde{f}$ is irreducible. All cyclotomic polynomials with exception of
  $\Phi_1 = x - 1$ are palindromes. $\Phi_1$ does not divide $f$ because
  $f(1) = \sum\limits_{k=0}^{d-1} f_k \ge 2$. Therefore, $f$ has at most one
  non-palindrome factor and hence $A$ does not have non-trivial collisions.
\end{proof}

For the multidimensional case, our bound is more pessimistic:
\begin{lemma}
  Let $A$ be a sufficiently big random $D-dimensional$ array of 0's and 1's
  where each element is chosen randomly and independently with equal probability
  $1/2$ with exception of the first and the last elements found on the edge
  $\mathcal{L}$, which are 1's. Then, with probability at least
  \begin{equation*}
    p = 1 - \sqrt{\frac{2}{\pi d}}
  \end{equation*}
  where $d+1$ is the length of $\mathcal{L}$, $A$ does not have non-trivial
  collisions in the two-point correlation function.
\end{lemma}
\begin{proof}
  Firstly, assume $D = 2$. Let $f(x_1, x_2) \in \mathbb{Z}[x_1][x_2]$ be an
  image of $A$ under Z-transform and change of variables $x_i = z_i^{-1}$. Let
  $\phi_l$ be a mapping $\mathbb{Z}[x_1][x_2] \rightarrow \mathbb{Z}[x_1]$ which
  ``extracts'' the leading coefficient $f_l \in \mathbb{Z}[x_1]$ from a
  bivariate polynomial $f$ and $\phi_f$ be a mapping
  $\mathbb{Z}[x_1][x_2] \rightarrow \mathbb{Z}[x_1]$ which ``extracts'' the free
  coefficient $f_0 \in \mathbb{Z}[x_1]$.

  $\phi_f$ and $\phi_l$ are homomorphisms from multiplicative monoid
  $\mathbb{Z}^*[x_1][x_2]$ to $\mathbb{Z}^*[x_1]$.
  Indeed if $g = g_0 + g_1x_2 + \dots + g_nx_2^n$ and
  $h = h_0 + h_1x_2 + \dots + h_mx_2^m$ then
  \begin{align*}
    \phi_f(g h) &= \phi_f[(g_0  h_0) + (g_0 h_1 + g_1 h_0)x_2 + \dots + (g_n
      h_m) x_2^{n+m}] \\
    &= g_0 h_0 = \phi_f(g) \phi_f(h) \qquad \text{and} \\
    \phi_l(g h) &= \phi_l[(g_0  h_0) + (g_0 h_1 + g_1 h_0)x_2 + \dots + (g_n
      h_m) x_2^{n+m}] \\
    &= g_n h_m = \phi_l(g) \phi_l(h)
  \end{align*}

  Now, without loss of generality, assume that the free (resp. leading)
  coefficient $f_k$ of $f$ has a degree $d$ and a non-zero free coefficient. If
  it is irreducible in $\mathbb{Z}[x_1]$, then $f$ is irreducible in
  $\mathbb{Z}[x_1][x_2]$. Indeed, if $f$ is not irreducible, then $f = p_1 p_2$
  for some non-constant $p_1, p_2$. Applying $\phi = \phi_f$ (resp. $\phi_l$) to
  this equality, we get $\phi(f) = f_k = \phi(p_1) \phi(p_2)$ for non-constant
  $\phi(p_1), \phi(p_2)$ which contradicts with $f_k$ being irreducible. Using
  \cref{eq:prob-irr} we obtain the least bound for the probability of $f$ being
  irreducible. If $f$ is irreducible, then it cannot have two or more
  non-palindrome factors and therefore cannot have non-trivial collisions, which
  proves the lemma.

  The proof for $D > 2$ is by recursion.
\end{proof}

Despite the probability of ``meeting'' a random polynomial which has non-trivial
collisions in the two-point function is decreasing with increase of its degree,
we can easily construct ``subtrivial'' collisions of an arbitrary (sufficiently
big) degree. Let's take two polynomials $f_1, f_2$ with coefficients being in
$\{0, 1\}$. Obviously, a polynomial $f(x) = f_1(x) f_2(x^{\deg f_1 + k})$ with
integer $k > 0$ also has its coefficients in $\{0, 1\}$.

\begin{example}
  Two sequences
  \begin{align*}
    & 1101110100001101 \quad \text{and} \\
    & 1101000011011101
  \end{align*}
  are two ``subtrivial'' collisions obtained by choosing $f_1 = f_2 = 1 + x + x^3$.
\end{example}

\section{Results}

\bibliographystyle{elsarticle-num}
\bibliography{paper}
\end{document}